---
title: "Moter Trend -  The relationship between a set of variables and miles per gallon"
author: "Tom Lous"
date: "26 April 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

We are looking at a data set of a collection of cars, and are interested in exploring the relationship between a set of variables and miles per gallon. In particularly the following two questions:

* “Is an automatic or manual transmission better for MPG”
* "Quantify the MPG difference between automatic and manual transmissions"


## Processing

First change 'am' to  factor (0 = automatic, 1 = manual)
Also use this for other properties: 

```{r}
data(mtcars)

mtcars$am <- as.factor(mtcars$am)
levels(mtcars$am) <- c("automatic", "manual")

mtcars$cyl <- as.factor(mtcars$cyl)



```

## Exploratory data analyses 

Look at the ranges of 
```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(GGally)
library(dplyr)
p <- ggplot(mtcars, aes(am, mpg))
p + geom_boxplot(aes(fill = am))

mtcars %>% 
   select(mpg, disp, hp, wt, cyl, am) %>% 
   ggpairs(
     mapping = ggplot2::aes(color = am), 
     upper = list(continuous = "density", combo = "box"), 
     lower = list(continuous = "smooth"))
```

